% 
% EMSEC seminar template, 24.09.2010
% david.oswald@rub.de
% 
% based on
% EMSEC thesis template, 13.09.2010
% 
% Calls all chapters
% Each chapter is contained in a seperate "\input" file 
% Style: doublesided, scrreport, dina4
%
% Benedikt Driessen, 2010
% benedikt.driessen@rub.de
%

% *************************************************************
% ENTER INFORMATION ABOUT AUTHOR, TITLE and TYPE HERE
\newcommand{\thauthor}{Julian Speith, Felix Haarmann}
\newcommand{\thtitle}{Side-Channel Attacks on Implementations of Lattice-Based Cryptosystems}
% *************************************************************

\documentclass[a4paper,12pt,twoside,openany,headsepline,bibliography=totocnumbered]{scrbook}

% Import settings, packages, etc.
\input{settings}

\begin{document}

% Import frontpage
\input{frontpage}

% Abstract
\section*{Abstract}
Due to the progress in the developement of quantum computers, the need for pratical post-quantum cryptography is getting increrasingly stronger. Lattice-based cryptography is just one of a few candidates that could replace the public-key algorithms used these days. However, it is the most promising one.

Though lattice-based cryptography has only come up in recent years, a lot of research has been done in that area. Most papers in recent years have focused on efficient implementations for some of the proposed lattice-based algorithms, without considering the possibility of side-channel attacks. Therefore, we summarize a few proposed side-channel attacks and  countermeasures and evaluate their practicability and efficiency.

In that context, we present two masking schemes and their implementation for a ring-LWE encryption scheme. While just one of whom is sound against all first-order DPA attacks, it lacks in efficiency compared to the other one. The second one is much more efficient, but might still be vulnerable to some refined first-order DPA.

Furthermore, we describe a cache attack on the Gaussian sampling of the Bimodal Lattice Signature Scheme (BLISS) by introducing two different sampling algorithms and explain cache attacks on both. Afterwards, we present experimental attack results with a perfect side channel and a real side channel on Intel CPUs.

Finally, two blinding techniques are shown, which can be used for any arbitrary algorithm, but are not proven to be secure against side-channel attacks. However, they do not add too many computations to the existing algorithms, meaning there will not be a big loss in terms of efficiency.
\clearpage

\tableofcontents
\mainmatter

% List of acronyms
\chapter*{Acronyms}
\begin{acronym}
 	\setlength{\itemsep}{0.2em} 
	\acro{BLISS}{Bimodal Lattice Signature Scheme}
 	\acro{DPA}{Differential Power Analysis}
 	\acro{HO-DPA}{Higher Order Differential Power Analysis}
 	\acro{LPR}{Lyubashevsky-Peikert-Regev}
 	\acro{NTT}{Number Theoretic Transform}
 	\acro{PRNG}{Pseudo Random Number Generator}
 	\acro{ring-LWE}{Ring Learning with Errors Problem}
\end{acronym}

%
% Include all your chapters as .tex files, each file contains sections \section{name of section},
% subsections \subsections{...} and so on...
%

\pagenumbering{arabic} 

\input{01_introduction}
\input{02_background}
\input{03_masked1}
\input{04_masked2}
\input{05_bliss}
\input{06_blinding}
\input{07_conclusion}

% Generate list of figures
\newpage
\listoffigures

% Generate list of tables
\newpage
\listoftables

% Generate list of algorithms
\newpage
\listofalgorithms
\clearpage

% Generate bibliography with bibtex, the bibfile here is "bibliography.bib", use alphanumerical style
\begin{appendix}
 	\bibliography{bibliography}
 	\bibliographystyle{alpha}
\end{appendix}

\end{document} 
